Demystifying the Computer Science \\ PhD Admission in the US\\{\large A Handbook for International \emph{and Domestic} Students

### 2024-07-20 demystify.tex
services

          \item Increasing admission chance by being unique and standing out (\autoref{sec:improve-your-chance}).
          \item It take about 10 minutes to review an application (\autoref{sec:ievaluate}).
          \item Being a minority can help (\autoref{sec:urm}).
          \item You can successfully apply to CS PhD even if you have non-CS background (\autoref{sec:non-stem}).
          \item You do not need an MS (\autoref{sec:msrequirement}), and CS PhD takes about 5--7 years (\autoref{sec:time}). Compared to other countries, CS PhD in the US does not require an MS degree but has a longer PhD study time (\autoref{sec:non-us-differences}).
          \item What should you call your profs.? (\autoref{sec:address}) Many possibilities (e.g., prof., Dr., and even their firstnames), but probably not Mr. or Mrs.
          \item Your entire PhD program costs about \$400K in total, but you \emph{do not} pay for it (\autoref{sec:ra-cost})
          \item Ask your prof. if they can buy computer equipment and such for your research (\autoref{sec:buying-equipment}).
          %\item Despite some miserable stories on social media, many PhD students have good mentors, supportive lab mates, healthy working environment ... and are happy (\autoref{sec:happy}).
        \end{itemize}

@@

On average, I spend about 10 minutes reviewing each application (less for clear rejections and more for potential acceptances). While this seems short, it is not that difficult to tell if an application is good or bad.  In fact, this time
is twice what other faculty spend on average, e.g., Philip Guo spent \href{https://pg.ucsd.edu/PhD-application-tips.htm}{3--5 minutes} per application.

@@

>>>>

          \item Increasing admission chance by being unique and standing out (\autoref{sec:improve-your-chance}).
          \item It takes about 10 minutes to review an application (\autoref{sec:ievaluate}).
          \item Sometimes adcom members do dicuss applicants, but in most cases they make independent decisions (\autoref{sec:adcom-discuss}).           
          \item Being a minority can help (\autoref{sec:urm}).
          \item You can successfully apply to CS PhD even if you have non-CS background (\autoref{sec:non-stem}).
          \item You do not need an MS (\autoref{sec:msrequirement}), and CS PhD takes about 5--7 years (\autoref{sec:time}). Compared to other countries, CS PhD in the US does not require an MS degree but has a longer PhD study time (\autoref{sec:non-us-differences}).
          \item What should you call your profs.? (\autoref{sec:address}) Many possibilities (e.g., prof., Dr., and even their firstnames), but probably not Mr. or Mrs.
          \item Your entire PhD program costs about \$400K in total, but you \emph{do not} pay for it (\autoref{sec:ra-cost})
          \item Ask your prof. if they can buy computer equipment and such for your research (\autoref{sec:buying-equipment}).
          %\item Despite some miserable stories on social media, many PhD students have good mentors, supportive lab mates, healthy working environment ... and are happy (\autoref{sec:happy}).
        \end{itemize}

@@

On average, I spend about 10 minutes reviewing each application (less for clear rejections and more for potential acceptances). While this seems short, it is not that difficult to tell if an application is good or bad.  In fact, this
is twice what other faculty spend on average, e.g., Philip Guo spent \href{https://pg.ucsd.edu/PhD-application-tips.htm}{3--5 minutes} per application.

@@

\section{Do adcom members talk to each other?}\label{sec:adcom-discuss}

In most cases we make decision independently and do not talk to each other.
However, we do discuss when there are discrepancies in evaluations, in which case the adcom chair will ask reviewers discuss the application to reach a consensus.  We also chat to each other for interesting or strong applications that we think should be admitted, e.g., how to recruit this student or who should be the adviser. If the student mentioned a faculty member in their SOP, we might ask that faculty member if they are interested in the student.  In short, while we do not talk to each other for most applications, we do for some.
